Edge sign prediction in general graphs and connection with Correlation Clustering

### 2017-06-30 gtx_algorithm.tex
more text on the singleton problem

While all $G_t$ are connected, it might happen during the execution of \extractStar{} that%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_singleton_tikz.pdf}
  \caption{The formation of a singleton star}
  \label{fig:gtx_singleton}
\end{marginfigure}
the degree of a node $u$ drops to zero because all of its neighbors were claimed by the
periphery of previous stars. Such a node then forms a \emph{singleton star}, as illustrated in
\autoref{fig:gtx_singleton}, where after the creation of \starone{1} and \starone{2}, $c_3$ is the
single node of the third star \starone{3}.%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_noreduc_tikz.pdf}
  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
  \label{fig:gtx_noreduc}
\end{marginfigure}
In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
reduced by at least $2$ (because each star is made of at least two nodes). On the other hand, we can
construct a graph with singletons where the factor of reduction in number of nodes and edges can be
arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc} and assume that we first extract two stars
centered in $c_1$ and $c_2$.%

>>>>

While all $G_t$ are connected, it might happen during the execution of \extractStar{} that%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_singleton_tikz.pdf}
  \caption{The formation of a singleton star}
  \label{fig:gtx_singleton}
\end{marginfigure}
the degree of a node $u$ drops to zero because all of its neighbors were claimed by the
periphery of previous stars. Such a node then forms a \emph{singleton star}, as illustrated in
\autoref{fig:gtx_singleton}, where after the creation of \starone{1} and \starone{2}, $c_3$ is the
single node of the third star \starone{3}.%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.95\linewidth]{assets/tikz/gtx_noreduc_tikz.pdf}
  \caption{A case were too many singletons \enquote{waste} one iteration of \gtx{}}
  \label{fig:gtx_noreduc}
\end{marginfigure}
In the case no singleton are formed during one execution of \extractStar{}, the number of nodes is
reduced by at least a factor $2$ (because each star is made of at least two nodes). On the other hand, we can
construct a graph with singletons where the factor of reduction in number of nodes and edges can be
arbitrarily close to $1$. Consider \autoref{fig:gtx_noreduc}, where $c_1$ and $c_2$ have a degree
$p$ and their neighbors (in light gray) all have a degree $q$, which we can vary from $1$ to $p-1$.
At this step $t$, we thus have $|V_t| = 2(1+p)+pq$ and $|E_t| = 2(p+qp)$. Next we run \extractStar{}
and assume that we first extract the two stars centered in $c_1$ and $c_2$. This reduces the
effective degree of all the middle nodes (in dark gray) to $0$ and they thus form singletons, that
are all connected to the first two stars by \collapseStar{}. As a result, at step $t+1$ we have
$|V_{t+1}| = 2 + pq$ and $|E_{t+1}| = 2pq$, so that
\begin{equation*}
\frac{|V_{t+1}|}{|V_t|} = \frac{pq+2}{p(q+2) + 2} \qquad \text{and} \qquad
\frac{|E_{t+1}|}{|E_t|} = \frac{2pq}{2p(q+1)} = \frac{q}{q+1}
\end{equation*}
When we let $p$ goes to infinity, setting $q=1$ makes $\frac{|V_{t+1}|}{|V_t|}$ tend to
$\nicefrac{1}{3}$ and $\frac{|E_{t+1}|}{|E_t|}$ to $\nicefrac{1}{2}$ while setting $q=p-1$ makes
both $\frac{|V_{t+1}|}{|V_t|}$ and $\frac{|E_{t+1}|}{|E_t|}$ tend to $1$. Note however that in both
cases, this is not really a problem because at the next iteration, there will be only two stars (one
of them will be a singleton). Is it always like that? In other words, can we have a situation where
a large number of singletons is carried over several consecutive iterations?
